Meeting the Challenges of Modeling Astrophysical Thermonuclear Explosions: \castro, \maestro, and the \amrex\ Astrophysics Suite

### 2017-10-24 astronum_2017/astronum-amrex.tex
more updates

Our latest focus has been on GPU ports of our application codes.  A
small proxy app, \starlord, was created from \castro\ with just the
hydrodynamics and stellar equation of state.  It uses a simple
method-of-lines formulation of the hydrodynamics and advects 13
nuclear species in addition to the hydrodynamics.  To offload work on
GPUs, GPU support was added directly into \amrex\footnote{This support
  is currently on a feature branch in the git repo, awaiting merge
  into {\tt development}.}.  In \amrex, grids consist of a number of
boxes that enable domain decomposition for parallelism.\MarginPar{unclear what the difference between a grid and box is here}
An iterator loops over all of the boxes at the same level of refinement and passes
a data pointer into a Fortran kernel function where the work is
done. For the simulation state data that resides in each box, we have
modified the memory allocator so that it can use a CUDA allocator
(either device or managed memory).  The domain iterator is configured
to handle data motion to and from the device, so that the compute
kernels can operate on data that is presumed to already be there, and
the computation is decoupled from the memory management. Computation
on the data can then be performed with OpenACC, CUDA Fortran, or (more
recently) OpenMP 4.5. We have also built CUDA compute support into
\amrex\ so that a compute kernel can be transparently operated on
using CUDA Fortran without substantially modifying the kernels (and we
anticipate using a similar strategy to use OpenACC and/or OpenMP in
the future). This helps ensure that we can continue to maintain
performance portability in our simulation codes, by decoupling the
physics algorithms from the backend support used to implement them on
various compute architectures.\MarginPar{add CPU time for reference}

@@

Figure~\ref{fig:current-runs} shows some of our recent science
simulations.  The left panel is an image of convection in the helium
layer on a sub-Chandra white dwarf.  This is part of a study of the
early stages of the double detonation SNe Ia model.  Using \maestro,
we are able to model the convection in the He layer for many turnover
times and saw a range of outcomes depending on the mass of the white
dwarf and He layer, including a global He-nova-like behavior where the
entire layer runs way together and localized
burning~\cite{subchandra2}.\MarginPar{incomplete sentence}
We are performing further studies to characterize the ignition.

@@

\ack The work at Stony Brook was supported by DOE/Office of Nuclear
Physics grant DE-FG02-87ER40317 and NSF award AST-1211563.  The work
at LBNL was supported by the DOE Office of Advanced Scientific
Computing Research under Contract No, DE-AC02-05CH11231. An award of
computer time was provided by the Innovative and Novel Computational
Impact on Theory and Experiment (INCITE) program.  This research used
resources of the Oak Ridge Leadership Computing Facility at the Oak
Ridge National Laboratory, which is supported by the Office of Science
of the U.S. Department of Energy under Contract
No.\ DE-AC05-00OR22725.  We appreciate the efforts of the OLCF (and in
particular Fernanda Foertter) in organizing the GPU hackathons.  This
research used resources of the National Energy Research Scientific
Computing Center, which is supported by the Office of Science of the
U.S. Department of Energy under Contract No.\ DE-AC02-05CH11231.
Results in this paper were obtained using the high-performance LIred
computing system at the Institute for Advanced Computational Science
at Stony Brook University, which was obtained through the Empire State
Development grant NYS \#28451.  The GPU work benefited from a grant of
an Titan X Pascal board from Nvidia Corp through the GPU Grant
Program.  This research has made use of NASA's Astrophysics Data
System Bibliographic Services.

>>>>

Our latest focus has been on GPU ports of our application codes.  A
small proxy app, \starlord, was created from \castro\ with just the
hydrodynamics and stellar equation of state.  It uses a simple
method-of-lines formulation of the hydrodynamics and advects 13
nuclear species in addition to the hydrodynamics.  To offload work on
GPUs, GPU support was added directly into \amrex\footnote{This support
  is currently on a feature branch in the git repo, awaiting merge
  into {\tt development}.}.  In \amrex, an iterator loops over all of
the boxes at the same level of refinement and passes a data pointer
into a Fortran kernel function where the work is done. For the
simulation state data that resides in each box, we have modified the
memory allocator so that it can use a CUDA allocator (either device or
managed memory).  The domain iterator is configured to handle data
motion to and from the device, so that the compute kernels can operate
on data that is presumed to already be there, and the computation is
decoupled from the memory management. Computation on the data can then
be performed with OpenACC, CUDA Fortran, or (more recently) OpenMP
4.5. We have also built CUDA compute support into \amrex\ so that a
compute kernel can be transparently operated on using CUDA Fortran
without substantially modifying the kernels (and we anticipate using a
similar strategy to use OpenACC and/or OpenMP in the future). This
helps ensure that we can continue to maintain performance portability
in our simulation codes, by decoupling the physics algorithms from the
backend support used to implement them on various compute
architectures.\MarginPar{add CPU time for reference}

@@

Figure~\ref{fig:current-runs} shows some of our recent science
simulations.  The left panel is an image of convection in the helium
layer on a sub-Chandra white dwarf.  This is part of a study of the
early stages of the double detonation SNe Ia model.  Using \maestro,
we are able to model the convection in the He layer for many turnover
times and saw a range of outcomes depending on the mass of the white
dwarf and He layer, including a both nova-like behavior where the
entire layer runs way together and localized
burning ignited in a small region~\cite{subchandra2}.
We are performing further studies to characterize the ignition.

@@

\ack The work at Stony Brook was supported by DOE/Office of Nuclear
Physics grant DE-FG02-87ER40317 and NSF award AST-1211563.  The work
at LBNL was supported by the DOE Office of Advanced Scientific
Computing Research under Contract No, DE-AC02-05CH11231. An award of
computer time was provided by the Innovative and Novel Computational
Impact on Theory and Experiment (INCITE) program.  This research used
resources of the Oak Ridge Leadership Computing Facility at the Oak
Ridge National Laboratory, which is supported by the Office of Science
of the U.S. Department of Energy under Contract
No.\ DE-AC05-00OR22725.  We appreciate the efforts of the OLCF (and in
particular Fernanda Foertter) in organizing the GPU hackathons.  This
research used resources of the National Energy Research Scientific
Computing Center, which is supported by the Office of Science of the
U.S. Department of Energy under Contract No.\ DE-AC02-05CH11231.
Results in this paper were obtained using the high-performance LIred
computing system at the Institute for Advanced Computational Science
at Stony Brook University, which was obtained through the Empire State
Development grant NYS \#28451.  The GPU work benefited from a grant of
an Titan X Pascal board from Nvidia Corp through the GPU Grant
Program.  Visualization were done using \cite{yt}.  This research has
made use of NASA's Astrophysics Data System Bibliographic Services.
